---
title: "An introduction to biodbKegg"
author: "Pierrick Roger"
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('biodbKegg')`"
vignette: |
  %\VignetteIndexEntry{Introduction to the biodbKegg package.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: false
  BiocStyle::pdf_document: default
---

# Purpose

biodbKegg is a *biodb* extension package that implements a connector to
KEGG Compound database.

# Initialization

The first step in using *biodbKegg*, is to create an instance of the biodb
class `Biodb` from the main *biodb* package. This is done by calling the
constructor of the class:
```{r, results='hide'}
mybiodb <- biodb::Biodb()
```
During this step the configuration is set up, the cache system is initialized
and extension packages are loaded.

We will see at the end of this vignette that the *biodb* instance needs to be
terminated with a call to the `terminate()` method.

# Creating a connector to KEGG Compound database

In *biodb* the connection to a database is handled by a connector instance that
you can get from the factory.
biodbKegg implements a connector to a remote database.
Here is the code to instantiate a connector:
```{r}
conn <- mybiodb$getFactory()$createConn('kegg.compound')
```

# Accessing entries

To retrieve entries, use:
```{r}
entries <- conn$getEntry(c('C00133', 'C00751'))
entries
```

To convert a list of entries into a dataframe, run:
```{r}
<<<<<<< HEAD
x <- mybiodb$entriesToDataframe(entries, compute=FALSE)
=======
x <- mybiodb$entriesToDataframe(entries)
>>>>>>> 2a061b6 (Upgrade with ExtPackage.)
x
```

# Search for compounds of a certain mass

```{r}
<<<<<<< HEAD
ids <- conn$searchForEntries(list(monoisotopic.mass=list(value=64, delta=2.0)), max.results=10)
=======
ids <- conn$searchCompound(mass = 64, mass.tol = 2.0, max.results = 10)
>>>>>>> 2a061b6 (Upgrade with ExtPackage.)
entries <- mybiodb$getFactory()$getEntry('kegg.compound', ids)
```


# Closing biodb instance

When done with your *biodb* instance you have to terminate it, in order to
ensure release of resources (file handles, database connection, etc):
```{r}
mybiodb$terminate()
```
